---
title: "chicago_eda"
author: "Linguo Ren"
date: "2024-11-24"
output: html_document
---

```{r}
## Load the Data
data_2013 <- read.csv("data/Crimes_2013-22.csv")
print(dim(data_2013))
```

```{r}
head(data_2013, 30)
```

```{r}
library(ggplot2)

# Grouping data by Year and calculating counts
df_count_per_year <- aggregate(counts ~ Year, data = transform(data_2013, counts = 1), FUN = sum)

# Print the data
print(df_count_per_year)

# Plotting the data using ggplot2
ggplot(df_count_per_year, aes(x = Year, y = counts)) +
  scale_x_continuous(breaks = seq(min(df_count_per_year$Year), max(df_count_per_year$Year), by = 1)) +
  geom_line(color="red")+
  labs(title = "Counts per Year", x = "Year", y = "Counts")
```

```{r}
## top 5 most common crime type in past 10 years, and only keep the top 10 most common ones
library(dplyr)
crime_type_list <- data_2013 %>% 
  group_by(Primary.Type) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>%
  head(5)
print(crime_type_list)

```

```{r, fig.width=20, fig.height=10}
# Load the data AND only crime_type_list for each year and plot it
top_5_crime_types <- data_2013 %>%
  filter(Primary.Type %in% crime_type_list$Primary.Type) %>%
  group_by(Year, Primary.Type) %>%
  summarise(counts = n(), .groups = "drop")

##plot it and include the * at each data time
ggplot(top_5_crime_types, aes(x = Year, y = counts, color = Primary.Type)) +
  geom_line() +
  geom_point(size = 1, shape = 16, color = "red") +
  scale_x_continuous(breaks = seq(min(top_5_crime_types$Year), max(top_5_crime_types$Year), by = 1)) +
  labs(
    title = "Top 5 Crime Types Over the Years",
    x = "Year",
    y = "Counts",
    color = "Primary Type"
  ) +
  theme_minimal()
```

```{r}
# Heatmap for top 5 primary types per year
ggplot(top_5_crime_types, aes(x = Primary.Type, y = Year , fill = counts)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "red") +
  labs(
    title = "Heatmap of Top 5 Crime Types Over the Years",
    x = "Primary Type",
    y = "year",
    fill = "Counts",
  ) +
  scale_y_continuous(breaks = seq(min(top_5_crime_types$Year), max(top_5_crime_types$Year), by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
## Check the total crime per district
crime_per_district <- data_2013 %>%
  filter(!is.na(District)) %>%
  group_by(District) %>%
  summarise(counts = n(), .groups = "drop") %>%
  arrange(District)

## Calcualte the arresting number
arrest_per_district <- data_2013 %>%
  filter(!is.na(District)) %>%
  filter(Arrest == "True") %>%
  group_by(District) %>%
  summarise(arrests = n(), .groups = "drop") %>%
  arrange(District)


## plot bar
ggplot(crime_per_district, aes(x = District, y = counts)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(
    title = "Total Crime per District",
    x = "District",
    y = "Counts"
  ) +
  scale_x_continuous(breaks = seq(min(crime_per_district$District), max(crime_per_district$District), by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# Filter for valid districts
valid_districts <- c(1:25)
data_2013 <- data_2013 %>%
  filter(District %in% valid_districts)

# Calculate total crimes per district
crime_per_district <- data_2013 %>%
  filter(!is.na(District)) %>%
  group_by(District) %>%
  summarise(counts = n(), .groups = "drop") %>%
  arrange(District)

# Calculate total arrests per district
arrest_per_district <- data_2013 %>%
  filter(!is.na(District)) %>%
  filter(Arrest == "True") %>%
  group_by(District) %>%
  summarise(arrests = n(), .groups = "drop") %>%
  arrange(District)

# Combine crimes and arrests into one data frame for plotting
crime_arrest_df <- left_join(
  crime_per_district,
  arrest_per_district,
  by = "District"
) %>%
  pivot_longer(
    cols = c(counts, arrests),
    names_to = "type",
    values_to = "counts"
  )

# Plot crimes and arrests on the same bar plot
ggplot(crime_arrest_df, aes(x = District, y = counts, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") + # Dodge for side-by-side bars
  labs(
    title = "Total Crimes and Arrests per District",
    x = "District",
    y = "Counts"
  ) +
  scale_fill_manual(
    values = c("counts" = "blue", "arrests" = "red"),
    name = "Type",
    labels = c("Arrests", "Crimes")
  ) +
  scale_x_continuous(breaks = seq(min(crime_per_district$District), max(crime_per_district$District), by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#### After creating the bar graph showing the total crimes per district in Chicago, I noticed some missing or invalid values in the district column. Upon further investigation, I discovered that these were not due to missing data but rather because certain districts, including 13, 21, and 23, are not actual police districts in Chicago. To ensure accuracy, I will remove these rows from the dataset and recreate the bar graph. For reference, information on Chicago’s official police districts can be found [here](https://www.chicagopolice.org/police-districts/).


```{r}
## Plot the horizontal line for arrest and counts based on the mean on orignal bar graph
library(dplyr)
library(ggplot2)

mean_crime_count <- mean(crime_arrest_df$counts[crime_arrest_df$type == "counts"])
mean_arrest_count <- mean(crime_arrest_df$counts[crime_arrest_df$type == "arrests"])

# Plot the bar chart with horizontal lines for mean values
ggplot(crime_arrest_df, aes(x = District, y = counts, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(
    yintercept = mean_crime_count, 
    linetype = "dashed",
    color = "blue"
  ) +
  geom_hline(
    yintercept = mean_arrest_count, 
    linetype = "dashed",
    color = "red"
  ) +
  annotate(
    "text", x = max(crime_arrest_df$District), y = mean_crime_count, 
    label = "Mean", color = "blue", vjust = -0.5, hjust = 1
  ) +
  annotate(
    "text", x = max(crime_arrest_df$District), y = mean_arrest_count, 
    label = "Mean", color = "red", vjust = -0.5, hjust = 1
  ) +
  labs(
    title = "Total Crimes and Arrests per District",
    x = "District",
    y = "Counts"
  ) +
  scale_fill_manual(
    values = c("counts" = "blue", "arrests" = "red"),
    name = "Type",
    labels = c("Arrests", "Crimes")
  ) +
  scale_x_continuous(breaks = seq(min(crime_arrest_df$District), max(crime_arrest_df$District), by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
library(tidyr)
crime_arrest_df <- cbind(
  crime_per_district %>% select(District, Crime = counts),
  arrest_per_district %>% select(Arrest = arrests) %>% mutate(Arrest = Arrest), 
  arrest_crime_ratio = arrest_per_district$arrests / crime_per_district$counts
)

## ggplot
ggplot(crime_arrest_df, aes(x = District, y = arrest_crime_ratio)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(
    title = "Arrest to Crime Ratio per District",
    x = "District",
    y = "Arrest to Crime Ratio"
  ) +
  scale_x_continuous(breaks = seq(min(crime_arrest_df$District), max(crime_arrest_df$District), by = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Assuming crime_arrest_df is a dataframe with columns "District", "arrest_crime_ratio", and "Arrest"

# Calculate the top 5 districts by arrest-crime ratio
top_5_arrest_ratio <- crime_arrest_df %>%
  arrange(desc(arrest_crime_ratio)) %>%
  head(5) %>%
  pull(District)

# Calculate the top 5 districts by number of arrests
top_5_crime_count <- crime_arrest_df %>%
  arrange(desc(Arrest)) %>%
  head(5) %>%
  pull(District)

print("Top 5 Districts by Arrest-Crime Ratio:")
print(top_5_arrest_ratio)

print("Top 5 Districts by Number of Crimes:")
print(top_5_crime_count)
```

```{r}
## using district one below as the line
south_district <- c(2,9,8,7,3,6,4,5,22)
north_district <- c(16,17,24,20,19,18)
west_district <- c(10,11,12,15,14,25)
center <- c(1)

# Function to calculate total values for a district group
calculate_district_summary <- function(districts, group_name, data) {
  data %>%
    filter(District %in% districts) %>%
    summarise(
      total_arrest_crime_ratio = sum(arrest_crime_ratio, na.rm = TRUE),
      total_arrest = sum(Arrest, na.rm = TRUE),
      total_crime = sum(Crime, na.rm = TRUE)
    ) %>%
    mutate(District_Group = group_name)
}

# Define the district groups
district_groups <- list(
  South = south_district,
  North = north_district,
  West = west_district,
  Center = center
)

# Apply the function to each district group and combine results
district_summary <- bind_rows(
  lapply(names(district_groups), function(group_name) {
    calculate_district_summary(district_groups[[group_name]], group_name, crime_arrest_df)
  })
)

# Plot the bar chart of ratios for the four district groups
ggplot(district_summary, aes(x = District_Group, y = total_arrest_crime_ratio)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(
    title = "Total Arrest-Crime Ratio per District Group",
    x = "District Group",
    y = "Total Arrest-Crime Ratio"
  )

```
```{r}
## 4 region total crime per year
# Define the function to calculate counts per year for a district group
calculate_counts_per_year <- function(data, districts, group_name) {
  data %>%
    filter(District %in% districts) %>%
    group_by(Year) %>%
    summarise(counts = n(), .groups = "drop") %>%
    mutate(District_Group = group_name)
}

# Define district groups
district_groups <- list(
  South = south_district,
  North = north_district,
  West = west_district,
  Center = center
)

# Apply the function to each district group and combine results
district_counts_per_year <- bind_rows(
  lapply(names(district_groups), function(group_name) {
    calculate_counts_per_year(data_2013, district_groups[[group_name]], group_name)
  })
)

# View the combined results
district_counts_per_year


## plot it
ggplot(district_counts_per_year, aes(x = Year, y = counts, color = District_Group)) +
  geom_line() +
  labs(
    title = "Total Crime Counts per Year by District Group",
    x = "Year",
    y = "Total Crime Counts",
  ) +
  scale_color_manual(
    values = c("South" = "red", "North" = "blue", "West" = "green", "Center" = "purple"),
    name = "District Group"
  ) +
  scale_x_continuous(breaks = seq(min(district_counts_per_year$Year), max(district_counts_per_year$Year), by = 1))

```

```{r}
## Plot the total arrest for each district
# Define the function to calculate total arrests per year for a district group
calculate_arrests_per_year <- function(data, districts, group_name) {
  data %>%
    filter(District %in% districts) %>%
    filter(Arrest == "True") %>%
    group_by(Year) %>%
    summarise(arrests = n(), .groups = "drop") %>%
    mutate(District_Group = group_name)
}

# Apply the function to each district group and combine results
district_arrests_per_year <- bind_rows(
  lapply(names(district_groups), function(group_name) {
    calculate_arrests_per_year(data_2013, district_groups[[group_name]], group_name)
  })
)

##plot it 
ggplot(district_arrests_per_year, aes(x = Year, y = arrests, color = District_Group)) +
  geom_line() +
  labs(
    title = "Total Arrest Counts per Year by District Group",
    x = "Year",
    y = "Total Arrest Counts",
  ) +
  scale_color_manual(
    values = c("South" = "red", "North" = "blue", "West" = "green", "Center" = "purple"),
    name = "District Group"
  ) +
  scale_x_continuous(breaks = seq(min(district_arrests_per_year$Year), max(district_arrests_per_year$Year), by = 1))
                     
```



```{r}
# Define a function to calculate arrest ratios per year
calculate_arrest_ratio_per_year <- function(data, district_group) {
  data %>%
    filter(District %in% district_group) %>%
    group_by(Year) %>%
    summarise(
      arrest_crime_ratio = sum(Arrest == "True") / n(),
      total_crime = n(),
      total_arrest = sum(Arrest == "True"),
      .groups = "drop"
    )
}

# South District
south_arrest_ratio_per_year <- calculate_arrest_ratio_per_year(data_2013, south_district)

# North District
north_arrest_ratio_per_year <- calculate_arrest_ratio_per_year(data_2013, north_district)

# West District
west_arrest_ratio_per_year <- calculate_arrest_ratio_per_year(data_2013, west_district)

# Center District
center_arrest_ratio_per_year <- calculate_arrest_ratio_per_year(data_2013, center)

# Combine them by columns and name each column with the region
arrest_ratio_per_year <- cbind(
  south_arrest_ratio_per_year %>% select(Year, South = arrest_crime_ratio),
  north_arrest_ratio_per_year %>% select(North = arrest_crime_ratio),
  west_arrest_ratio_per_year %>% select(West = arrest_crime_ratio),
  center_arrest_ratio_per_year %>% select(Center = arrest_crime_ratio)
)

# Plot line plot
ggplot(arrest_ratio_per_year, aes(x = Year)) +
  geom_line(aes(y = South, color = "South")) +
  geom_line(aes(y = North, color = "North")) +
  geom_line(aes(y = West, color = "West")) +
  geom_line(aes(y = Center, color = "Center")) +
  labs(
    title = "Arrest-Crime Ratio per Year by District Group",
    x = "Year",
    y = "Arrest-Crime Ratio",
    color = "District Group"
  ) +
  scale_x_continuous(breaks = seq(min(arrest_ratio_per_year$Year), max(arrest_ratio_per_year$Year), by = 1)) +
  scale_color_manual(
    values = c("South" = "red", "North" = "blue", "West" = "green", "Center" = "purple"),
    name = "District Group"
  )
```



